Fixed-Parameter Tractable Sampling for RNA Design with Multiple Target Structures

### 2019-01-25 BMC-Response-Round1.tex
spell check

\Comment{
Multi defect can be a good measure to understand the overall performance of the sequence designing method, but there is little explanation why Multidefect is enough to evaluate the quality of sequences for the multi (target-structure) objective problem in a practical use case such as Riboswitch design.
}%

@@

\Comment{
Although RNABluePrint was compared with MODENA in the previous study, RNABluePrint did not overwhelm MODENA in all of conditions (and it is unclear which version (variant) of RNABluePrint is applied in this study).
For that reason, RNARedPrint still should be compared with MODENA (in the same running time) using Multidefect and the criteria used in the previous studies such as $\delta$e1 and $\delta$e2. Also, RNARedprint and RNABlueprint are optimised to increase only a Multidefect value.}%

\Answer{As we tried to clarify in our general comments of this response letter (as well as by appropriate adaptions in the manuscript), RNAredprint does not by itself perform \emph{negative} design. It is a generation algorithm for (positively) designing sequences with complex properties. In this way, it addresses a large range of design use-cases, but is also able to generate good seed sequences, amenable to further optimization using other tools. For this reason, we focused our validation on this potential for better initialization, and compared the quality of sequences generated by RNARedPrint (Boltzmann weighted) to RNABlueprint (uniform). We noticed that, while RNAredprint is not explicitly tailored to optimize MultiDefect, it nevertheless produces sequences that: 1) achieve better values; and 2) are amenable to better further optimizations than the uniform version implemented by RNABluePrint.}

\Comment{Hence, the distribution of Multidefect, previous criteria, and the first and second term of Multidefect should be shown in box plots or violin plots to know the robustness of RNARedPrint. It is also interesting what is the characteristic of the sequence designed by each method in a specific example of functional RNA molecules.}%

@@

\Answer{
We share the concern of this reviewer that an unncecessarily imbalanced manuscript should be avoided, and moved less original/central parts of the method to the appendix. This includes the expression of various energy models as constraints, and the description of our naive numerical weight optimization. However, we want to emphasize that the techniques used in this paper are quite diverse, and stem from several subfields of computer science (combinatorial optimization, graph algorithms, enumerative/analytic combinatorics...). For this reason, we feel that the method part cannot be substantially shortened beyond the current point without making the text less accessible and poorly self-contained. 
}
%\AnswerTODO[Yann]{
%Move material to appendix
%Better motivations for \#P-hardness
%New results : 
%  1) Validation of stacking approximation: Correlation coefficient of Stacking and Turner (+ %goodness of fit)
%  2) Tree width and Time benchmark
%  3) Rejection analysis for multidim sampling
%  4) Proof-of-concept -Potential of better seeds
%}
\Comment{
2) Although, it was shown previously, that RNABluePrint is a leading tool for multiple design, it would be valuable to include and compare the results generated by the other tools, e.g. Frnakenstein.}%

@@

\Answer{
We understand that it is important to give an estimate of the actual performance and thus included computational times for the pre-computation, the Boltzmann sampling of 10\,000 sequences and the number of iterations for the multi-dimensional Boltzmann sampling for the complete benchmark set in Supl. Mat. Sections C and E.
}
\Comment{
4) The paper should be checked for English grammar and wording.}%

@@

\Answer{We were not completely sure what was meant by this comment. Just to be precise, we distinguish the term 'Boltzmann sampling' from 'multi-dimensional Bolzmann sampling' in our paper. 
On the one hand, Boltzmann sampling, i.e. drawing sequences from the Boltzmann distribution does not involve any rejection, and is thus exactly as fast as drawing them uniformly. For uniform sampling, RNARedPrint is usually even faster than RNABluePrint, due to the state-of-the-art tree decomposition underlying RNARedPrint, which typically yields lower treewidths than the graph decomposition of RNABluePrint.
On the other hand, multi-dimensional Boltzmann sampling requires a rejection step to target specific properties that are typically difficult to obtain by chance in the uniform distribution. There is, of course, an added cost associated to this rejection, and the simultaneous weight optimization. We do now provide detailed overview of the seed generation procedure, which like our multi-dimensional Boltzmann sampling strategy---runs through several iterations of performing the elementary Boltzmann sampling in Suppl.\ Mat. E. The corresponding Figure (Supp. Fig. 9) shows that typically only few iterations suffice to produce the sequences. In this way, the time for multi-dimensional Boltzmann sampling typically is a small multiple of the time required for uniform sampling.}

>>>>

\Comment{
Multi defect can be a good measure to understand the overall performance of the sequence designing method, but there is little explanation why MultiDefect is enough to evaluate the quality of sequences for the multi (target-structure) objective problem in a practical use case such as Riboswitch design.
}%

@@

\Comment{
Although RNABluePrint was compared with MODENA in the previous study, RNABluePrint did not overwhelm MODENA in all of conditions (and it is unclear which version (variant) of RNABluePrint is applied in this study).
For that reason, RNARedPrint still should be compared with MODENA (in the same running time) using MultiDefect and the criteria used in the previous studies such as $\delta$e1 and $\delta$e2. Also, RNARedPrint and RNABlueprint are optimized to increase only a MultiDefect value.}%

\Answer{As we tried to clarify in our general comments of this response letter (as well as by appropriate adaptions in the manuscript), RNARedPrint does not by itself perform \emph{negative} design. It is a generation algorithm for (positively) designing sequences with complex properties. In this way, it addresses a large range of design use-cases, but is also able to generate good seed sequences, amenable to further optimization using other tools. For this reason, we focused our validation on this potential for better initialization, and compared the quality of sequences generated by RNARedPrint (Boltzmann weighted) to RNABlueprint (uniform). We noticed that, while RNARedPrint is not explicitly tailored to optimize MultiDefect, it nevertheless produces sequences that: 1) achieve better values; and 2) are amenable to better further optimizations than the uniform version implemented by RNABluePrint.}

\Comment{Hence, the distribution of MultiDefect, previous criteria, and the first and second term of MultiDefect should be shown in box plots or violin plots to know the robustness of RNARedPrint. It is also interesting what is the characteristic of the sequence designed by each method in a specific example of functional RNA molecules.}%

@@

\Answer{
We share the concern of this reviewer that an unnecessarily imbalanced manuscript should be avoided, and moved less original/central parts of the method to the appendix. This includes the expression of various energy models as constraints, and the description of our naive numerical weight optimization. However, we want to emphasize that the techniques used in this paper are quite diverse, and stem from several subfields of computer science (combinatorial optimization, graph algorithms, enumerative/analytic combinatorics...). For this reason, we feel that the method part cannot be substantially shortened beyond the current point without making the text less accessible and poorly self-contained. 
}
%\AnswerTODO[Yann]{
%Move material to appendix
%Better motivations for \#P-hardness
%New results : 
%  1) Validation of stacking approximation: Correlation coefficient of Stacking and Turner (+ %goodness of fit)
%  2) Tree width and Time benchmark
%  3) Rejection analysis for multidim sampling
%  4) Proof-of-concept -Potential of better seeds
%}
\Comment{
2) Although, it was shown previously, that RNABluePrint is a leading tool for multiple design, it would be valuable to include and compare the results generated by the other tools, e.g. Frnakenstein.}%

@@

\Answer{
We understand that it is important to give an estimate of the actual performance and thus included computational times for the pre-computation, the Boltzmann sampling of 10\,000 sequences and the number of iterations for the multi-dimensional Boltzmann sampling for the complete benchmark set in Suppl. Mat. Sections C and E.
}
\Comment{
4) The paper should be checked for English grammar and wording.}%

@@

\Answer{We were not completely sure what was meant by this comment. Just to be precise, we distinguish the term 'Boltzmann sampling' from 'multi-dimensional Boltzmann sampling' in our paper. 
On the one hand, Boltzmann sampling, i.e. drawing sequences from the Boltzmann distribution does not involve any rejection, and is thus exactly as fast as drawing them uniformly. For uniform sampling, RNARedPrint is usually even faster than RNABluePrint, due to the state-of-the-art tree decomposition underlying RNARedPrint, which typically yields lower treewidths than the graph decomposition of RNABluePrint.
On the other hand, multi-dimensional Boltzmann sampling requires a rejection step to target specific properties that are typically difficult to obtain by chance in the uniform distribution. There is, of course, an added cost associated to this rejection, and the simultaneous weight optimization. We do now provide detailed overview of the seed generation procedure, which like our multi-dimensional Boltzmann sampling strategy---runs through several iterations of performing the elementary Boltzmann sampling in Suppl.\ Mat. E. The corresponding Figure (Supp. Fig. 9) shows that typically only few iterations suffice to produce the sequences. In this way, the time for multi-dimensional Boltzmann sampling typically is a small multiple of the time required for uniform sampling.}

### 2019-01-25 RNARedPrint.tex
spell check

\begin{abstractbox}
\begin{abstract}
  \parttitle{Background}
  The design of multi-stable RNA molecules has important applications in biology, medicine, and biotechnology. Synthetic design approaches profit strongly from effective in-silico methods, \revised{which substantially reduce the need for costly wet-lab experiments.} \\
%
  \parttitle{Results} We devise a novel approach to a central ingredient of most in-silico
  design methods: the generation of sequences \revised{that fold well into multiple target structures.} 
  % 
  \revised{Based on constraint networks, our approach \ourprog{}} supports generic
  Boltzmann-weighted sampling, which enables the positive design of RNA sequences
  with specific free energies (for each of multiple, possibly pseudoknotted, target structures) and \GCb-content.
%
  Moreover, we study general properties of our approach empirically and generate
  biologically relevant multi-target Boltzmann-weighted designs for a
  \revised{established} design benchmark. \revised{Our results demonstrate the effectivity and feasibility of the method in practice
  as well as the benefits of Boltzmann sampling over the previously best multi-target sampling strategy---even for the case of negative design of multi-stable RNAs.}
%
  \revised{Besides empirically studies, we finally justify the algorithmic details due to a fundamental theoretic result about multi-stable RNA design, namely the \#P-hardness of the counting of designs.}
%
  \parttitle{Conclusion} \ourprog{} introduces a novel, flexible and---as we demonstrate---effective approach to multi-target RNA design, which promises broad applicability and extensibility.

@@

Moreover, to ensure a uniform algorithmic treatment of contributions, it is convenient to encode the weight $\pi$ of each feature in its contributions. This transformation works by multiplying all contributions with $ln(\pi)$, where $\pi$ is the weight of the corresponding feature, since then
$exp(-\ln(\pi)f(S))=\pi^{-f(S)}.$

@@

\revised{In this way, our framework enables new possibilities
in the field of RNA sequence design.}
%allowing to enforce additional
%constraints, like \GCb\%, while controlling the energy of
%multiple target structures.
%Thus, it presents a major advance over
%previously applied ad-hoc sampling and even efficient %uniform sampling
%strategies. 
%
%perspectives:
\revised{As particular advantage over previous sequence generation methods,} it is extensible to include various more complex sequence constraints, including mandatory/forbidden motifs at specific
positions or anywhere in the designed sequences, by adapting formal language constructs
of Zhou \emph{et al}~\citet{Zhou2013}. In future work, negative design principles could 
be explicitly supported at the generation stage, for instance by
penalizing a set of alternative helices/structures.
%
\revised{We moreover envision using positive design to assess the significance of observed properties. Critically, our current perception of statisitical significance suffers from overly simplistic simple null models (\emph{e.g.} dinucleotide shuffling) used to model random RNAs~\cite{Rivas2017}). Here, our approach promises fundamental improvements by constructing null models of random sequences that satisfy multiple complex constraints.}

>>>>

\begin{abstractbox}
\begin{abstract}
  \parttitle{Background}
  The design of multi-stable RNA molecules has important applications in biology, medicine, and biotechnology. Synthetic design approaches profit strongly from effective in-silico methods, \revised{which substantially reduce the need for costly wet-lab experiments.} \\
%
  \parttitle{Results} We devise a novel approach to a central ingredient of most in-silico
  design methods: the generation of sequences \revised{that fold well into multiple target structures.} 
  % 
  \revised{Based on constraint networks, our approach \ourprog{}} supports generic
  Boltzmann-weighted sampling, which enables the positive design of RNA sequences
  with specific free energies (for each of multiple, possibly pseudoknotted, target structures) and \GCb-content.
%
  Moreover, we study general properties of our approach empirically and generate
  biologically relevant multi-target Boltzmann-weighted designs for a
  \revised{established} design benchmark. \revised{Our results demonstrate the efficacy and feasibility of the method in practice
  as well as the benefits of Boltzmann sampling over the previously best multi-target sampling strategy---even for the case of negative design of multi-stable RNAs.}
%
  \revised{Besides empirically studies, we finally justify the algorithmic details due to a fundamental theoretic result about multi-stable RNA design, namely the \#P-hardness of the counting of designs.}
%
  \parttitle{Conclusion} \ourprog{} introduces a novel, flexible and---as we demonstrate---effective approach to multi-target RNA design, which promises broad applicability and extensibility.

@@

Moreover, to ensure a uniform algorithmic treatment of contributions, it is convenient to encode the weight $\pi$ of each feature in its contributions. This transformation works by multiplying all contributions with $\ln(\pi)$, where $\pi$ is the weight of the corresponding feature, since then
$exp(-\ln(\pi)f(S))=\pi^{-f(S)}.$

@@

\revised{In this way, our framework enables new possibilities
in the field of RNA sequence design.}
%allowing to enforce additional
%constraints, like \GCb\%, while controlling the energy of
%multiple target structures.
%Thus, it presents a major advance over
%previously applied ad-hoc sampling and even efficient %uniform sampling
%strategies. 
%
%perspectives:
\revised{As particular advantage over previous sequence generation methods,} it is extensible to include various more complex sequence constraints, including mandatory/forbidden motifs at specific
positions or anywhere in the designed sequences, by adapting formal language constructs
of Zhou \emph{et al}~\citet{Zhou2013}. In future work, negative design principles could 
be explicitly supported at the generation stage, for instance by
penalizing a set of alternative helices/structures.
%
\revised{We moreover envision using positive design to assess the significance of observed properties. Critically, our current perception of statistical significance suffers from overly simplistic simple null models (\emph{e.g.} dinucleotide shuffling) used to model random RNAs~\cite{Rivas2017}). Here, our approach promises fundamental improvements by constructing null models of random sequences that satisfy multiple complex constraints.}
